---
output:
  word_document: default
  html_document: default
---

## Module 1 Assignment 2

#### Task 1
```{r}
library(tidyverse)
library(esquisse)
```

#### Task 2  
```{r}
diamonddata = diamonds
nrow(diamonddata)
ncol(diamonddata)
```

The number of rows in diamond data set is `r nrow(diamonddata)`.
The number of coumns in the diamond data set is `r ncol(diamonddata)`.

#### Task 3
```{r}
ggplot(diamonddata, aes(x = carat, y = price)) + geom_point()
```

#### Task 4
```{r}
ggplot(diamonddata, aes(x = carat, y = price, color = cut)) + geom_point()
```

#### Task 5
```{r}
ggplot(diamonddata, aes(x = carat, y = price, color = cut)) + geom_point() + facet_grid(color ~ .) 
```

From the visual above color and carat directly affect the cost and each other. Diamonds with the color D(best) have a higher price at a lower carat count. Whereas diamonds with the color J(Worst) need to have more carats to reach the price of the D diamonds. The cut doesn't seem to have much affect on the price since there's all types of cuts at the same price. The diamonds with the color D do seem to have better cuts, and vice versa more diamonds with the color J have worse cuts. 

#### Task 6
```{r}
InventoryData <- read_csv("InventoryData.csv")
```

#### Task 7
```{r}
inventoryA <- filter(InventoryData, Supplier == "A")
nrow(inventoryA)
```
inventoryA's data frame contains `r nrow(inventoryA)` rows.   

#### Task 8 
```{r}
inventoryA = mutate(inventoryA, OnHandRatio = `On Hand`/ `Annual Demand`)
```

The line of code above creates a new column in the inventoryA data table that divides the Annual Demand data by the On Hand data to give the On Hand ratio.

#### Task 9
```{r}
avg_costs <- InventoryData %>% group_by(`Item SKU`) %>% summarise(mean(`Cost per Unit ($)`))
```

#### Task 10
The part I found the most challenging was the facets, but reading up on it helped me understand pretty easily. 