---
output:
  word_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(tidymodels)
library(caret)
library(gridExtra)
library(vip)
library(ranger)
```

```{r}
drug <- read_csv("drug_data-1.csv")
```

```{r}
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
str(drug)
```

```{r}
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"
```

```{r}
drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
"45_54", "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels =
c("Under16", "At16", "At17", "At18", "SomeCollege",
"ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
  mutate(Country = factor(Country,
labels = c("USA", "NewZealand", "Other", "Australia",
"Ireland","Canada","UK"))) %>%
  mutate(Ethnicity = factor(Ethnicity,
labels = c("Black", "Asian", "White", "White/Black", "Other",
"White/Asian", "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
select(-ID)
```


```{r}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>%
  select(!(Semer:VSA))
names(drug_clean)
```

### Task 1

```{r}
summary(drug_clean)
```

There is no missingness in this dataset. 

### Task 2

```{r}
set.seed(1234)
drug_split = initial_split(drug_clean, prop=.70, strata = Nicotine)
train = training(drug_split)
test = testing(drug_split)

```

### Task 3

```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)
```
```{r}
p1 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
```
```{r}
p1 = ggplot(train, aes(x = Ascore, y = Nicotine)) + geom_boxplot()
p2 = ggplot(train, aes(x = Cscore, y = Nicotine)) + geom_boxplot()
p3 = ggplot(train, aes(x = Impulsive, y = Nicotine)) + geom_boxplot()
p4 = ggplot(train, aes(x = SS, y = Nicotine)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
```

The different scores, Asocre, Escore, etc. don't seem to have too much of an impact on nicotine consumption. The main indicators would be education, age, country, gender, SS, and impulsivness. With country seeming to be the largest indicator. 

### Task 4

```{r}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)

drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% 
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

rf_grid = grid_regular(
  mtry(range = c(2, 8)),
  min_n(range = c(5, 20)), 
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid 
)
  
```

```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")
```

### Task 5

```{r}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```
```{r}
final_rf_fit = fit(final_rf, train)
```
```{r}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

Sensation is the most important variable when it comes to nicotine consumption, followed by being aged 45-54 and having an openness to experience. 

### Task 6

```{r}
trainpredrf = predict(final_rf_fit, train)
head(trainpredrf)
```

```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```

```{r}
testpredrf = predict(final_rf_fit, test)
head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

The model does not perform well with the test set there is high degradation. There's a difference of about 21% from the training set to the testing set, showing their may be some over fitting with this model. 

### Task 7
This model could be used in healthcare and seeing indicators of disease. I wouldn't use this model due to the high degradation in change from the training set to the testing set. In healthcare maybe that would be okay saying someone is at a higher risk, than they really are but overall I still would be concerned with the accuracy of this model due to over fitting. 